---
title: "Script"
output: html_document
subtitle: "This is a subtitle"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose of the Workshop:

One of the main principles of scientific research is related to its reproducible quality.

>Reproducibility refers to the ability of a researcher to duplicate the results of a prior study using the same materials that were used by the original researcher (Goodman, Fanelli, and Ioannidis 2016). This is a cornerstone of scientific research because if we cannot reproduce previous studies, we cannot rely on their results, and we cannot build on them to generate new knowledge.



# Representing the data
Important data types in R

1. Clasess:
  - Numerical, Character, Integer or Logical
2. Objects
  - Vectors, Matrices, Data frames, Lists, Factors, Missing values
3. Operations
  - Subsetting, Logical subsetting
  
Source:   

## Classes
#### Numerical
```{r Var_numerique}
# Définition de Variable Numérique
a <- 1
b <- 2
class(a)
class(b)

# Opérations basiques
c <- a + b  # somme
c <- sum(a, b)# somme
c <- a - b  # rest

```

Lé résultat de la somme est `r c`

#### Character
```{r, Var_char}
# Définition d'une variable type 'character' (texte)
d <- "Bonjour"
e <- "Comment ça va?" 
#f <- d + e 
#class(a)

# Opérations basiques
f <- paste(d, e, sep = ", ")
```

#### Integer
```{r Var_int}
numberSons <- 10L
```

#### Logical

```{r}

```




## Objects
### Vectors
### Matrices
### Data frames
### List
### Factors
### Missing Values



### Import Data



# Data wrangling

1. You know the Five Main Verbs (5MV) of data wrangling
2. You will know the Pipe operator ('Ceci n'est pas une pipe')
3. Tidy data approach



# Final Exercise

## Gapminder challenge


- Inspired by the late, great [Hans Rosling](https://www.youtube.com/watch?v=jbkSRLYSojo)


Install and load the [`gapminder` data package](https://github.com/jennybc/gapminder)

```{r, eval=FALSE, echo=TRUE, include=TRUE}
install.packages("gapminder")
library(gapminder)
gapminder
```


Pick at least __two__ of the tasks below from the task menu and approach each with a table and figure. 

  * `dplyr` should be your main data manipulation tool
  * `ggplot2` should be your main visualization tool

## Tasks menu {-}

* The mean life expectancy across all years for France?

* Get the maximum and minimum of GDP per capita for all continents.

* Look at the spread of GDP per capita across countries within the continents.

* How does life expectancy vary across different continents? (Using a `ggplot`)




* Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy - you decide - a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.

* Make up your own! 


---


You're encouraged to reflect on what was hard/easy, problems you solved, helpful tutorials you read, etc.








 
